The Inconsistency in G\"odel's Ontological Argument: \\ A Success Story for AI in Metaphysics

### 2016-08-14 Papers/2014/Universalis/ModalCollapse.tex
added my DFG grant references
minor editing
Leibniz/Lenzen initial commit

%----------classification, keywords, date
\subjclass{
Prim. 03B15; % Higher-order logic and type theory
Sec. 68T15 % Theorem proving (deduction, resolution, etc.)
}
% \marginpar{Change?: 
% 03A02 % Philosophical aspects of logic and foundations
% 03B45 % Modal logic (including the logic of norms)
% 03B35 % Mechanization of proofs and logical operations
% 68T02 % Artificial Intelligence
% }

@@

\dedicatory{ }

@@

Nevertheless, G{\"o}del's work still leaves room for criticism.  In
particular, his axioms are so strong that, when assuming unrestricted
comprehension principles\footnote{A possible direction to remedy modal
  collapse, as studied e.g. by Koons
  \cite{koons06:_sobel_goedel_ontol_proof} is to impose restrictions
  on the domain of properties.}, 
%\marginpar{added footnote}
they entail a \emph{modal collapse}
\cite{Sobel1987,sobel2004logic}: everything that is the case is so
necessarily.  There has been an impressive body of recent and ongoing
work
(cf.~\cite{sobel2004logic,Fitting,anderson90:_some_emend_of_goedel_ontol_proof,AndersonGettings,bjordal99,fuhrmann05:_exist_notwen,Hajek2002,Hajek2008,ContemporaryBibliography}
and the references therein) proposing solutions for the modal
collapse.  The goal of this article is to discuss the modal
collapse from the point of view of the modal square of opposition. 
Ontological arguments typically rely on an inversion of the normal direction of entailment
in the modal square of opposition for one particular proposition (i.e. God's existence), 
and the modal collapse shows that this inversion in fact occurs for all propositions, resulting in a total collapse of the modal square of opposition.

%\marginpar{@Bruno: Shall we add a sentence or two here?}

>>>>

\thanks{This work has been supported by the German National Research Foundation (DFG) under grant BE 2501/9-1,2.}

%----------classification, keywords, date
\subjclass{
Prim. 03B15; % Higher-order logic and type theory
Sec. 68T15% Theorem proving (deduction, resolution, etc.)
}
% 03A02 % Philosophical aspects of logic and foundations
% 03B45 % Modal logic (including the logic of norms)
% 03B35 % Mechanization of proofs and logical operations
% 68T02 % Artificial Intelligence

@@

\dedicatory{}

@@

Nevertheless, G{\"o}del's work still leaves room for criticism.  In
particular, his axioms are so strong that, when assuming unrestricted
comprehension principles\footnote{A possible direction to remedy modal
  collapse, as studied e.g. by Koons
  \cite{koons06:_sobel_goedel_ontol_proof} is to impose restrictions
  on the domain of properties.}, 
they entail a \emph{modal collapse}
\cite{Sobel1987,sobel2004logic}: everything that is the case is so
necessarily.  There has been an impressive body of recent and ongoing
work
(cf.~\cite{sobel2004logic,Fitting,anderson90:_some_emend_of_goedel_ontol_proof,AndersonGettings,bjordal99,fuhrmann05:_exist_notwen,Hajek2002,Hajek2008,ContemporaryBibliography}
and the references therein) proposing solutions for the modal
collapse.  The goal of this article is to discuss the modal
collapse from the point of view of the modal square of opposition. 
Ontological arguments typically rely on an inversion of the normal direction of entailment
in the modal square of opposition for one particular proposition (i.e. God's existence), 
and the modal collapse shows that this inversion in fact occurs for all propositions, resulting in a total collapse of the modal square of opposition.

### 2016-08-14 Papers/2016/IJCAI/FinalVersion/main.tex
Leibniz/Lenzen initial commit

\documentclass{article}
\usepackage{ijcai15}

@@

@@

\usepackage{amsmath}
\usepackage{xspace}
\usepackage{modallogics}
\usepackage{graphicx,url}
\usepackage{txfonts} % needed for \Diamondblack
\usepackage{color}
\usepackage{xcolor}

@@

\begin{figure}[t]
\centerline{\includegraphics[width=1\columnwidth]{./Images/Scott_S5U.png}}
\caption{Full Automation of T3 in \SFiveU; Consistency of Scott's
  Axioms;  Automatic Verification of Modal Collapse} \label{Scott_S5U}
\end{figure}

@@

\subsection{Improved Embedding}\label{sec:improvedembedding}
\begin{figure}[t]
\centerline{\includegraphics[width=1\columnwidth,height=10cm]{./Images/QML_S5U.png}}
\caption{Improved Embedding of \SFiveU} \label{QML_S5U}
\end{figure}

@@

\begin{figure}
\centerline{\includegraphics[width=1\columnwidth]{./Images/UglyEssence.png}}
\caption{Definition of Essence using Old Syntax} \label{UglyEssence}
\end{figure}

@@

\subsection{Argument Reconstruction in Isabelle}  \label{sec:arg2}
\begin{figure}[t]
\centerline{\includegraphics[width=1\columnwidth]{./Images/InconsistencyIsabelleK.png}}
\caption{Inconsistency of G\"odel's Axioms in HOML \K verified in Isabelle/HOL} \label{InconsistencyIsabelleK}
\end{figure}
\begin{figure*}[t]
  \centering
  \begin{subfigure}[t]{0.715\textwidth}
    \includegraphics[width=\textwidth]{./Images/Manuscript2.png}
    \caption{G\"{o}del's manuscript, with mutually inconsistent axioms and
      definitions highlighted (with permission from the Kurt G\"odel Papers, Shelby White and Leon Levy Archives Center, Princeton, NJ, USA, on deposit at Princeton University)} \label{GoedelScript} 
  \end{subfigure}
   \begin{subfigure}[t]{0.28\textwidth}
     \includegraphics[width=\textwidth,height=7.2cm]{./Images/Inconsistency_S5U_direct.png}
     \caption{Inconsistency in HOML \SFiveU} \label{Inconsistency_S5U} 
   \end{subfigure}
% \includegraphics[width=.73\textwidth]{./Images/Manuscript2.png} \hfill
% \includegraphics[width=.26\textwidth]{./Images/Inconsistency_S5U_direct.png}
 \caption{The inconsistency in G\"{o}del's manuscript has been
   detected and verified by HOL ATPs} 
\end{figure*}
To verify the correctness of the informal argument explained above, it
was reconstructed in Isabelle/HOL, using Metis\footnote{Metis, unlike
  external provers such as \textsc{Leo-II} or Satallax, 
  constructs proofs in Isabelle's highly trusted kernel calculus.} to automate the
inessential parts (cf. Fig.~\ref {InconsistencyIsabelleK}). The essential use of the Empty Essence Lemma, on
the other hand, is explicitly stated, to ensure that Isabelle is
reconstructing the same argument. In fact, without the help of this
lemma, Metis is still not strong enough to refute G\"odel's
axioms.

>>>>

\documentclass{article}
\usepackage{ijcai16}

@@

\pdfinfo{
/Title (The Inconsistency in G\"odel's Ontological Argument: A Success Story for AI in Metaphysics)
/Author (Christoph Benzm\"uller, Bruno Woltzenlogel Paleo) }

@@

\usepackage{amsmath}
\usepackage{xspace}
\usepackage{graphicx,url}
\usepackage{txfonts} % needed for \Diamondblack
\usepackage{color}
\usepackage{xcolor}

%\usepackage{modallogics}
\newcommand{\logic}[1]{\textbf{#1}\xspace}
\newcommand{\KB}{\logic{KB}}
\newcommand{\KFour}{\logic{K4}}
\newcommand{\KFourB}{\logic{K4B}}
\newcommand{\K}{\logic{K}}
\newcommand{\KT}{\logic{KT}}
\newcommand{\SFour}{\logic{S4}}
\newcommand{\SFive}{\logic{S5}}
\newcommand{\SFiveU}{\logic{S5\textsuperscript{U}}}

@@

\begin{figure}[t]
\centerline{\includegraphics[width=1\columnwidth]{./Scott_S5U.png}}
\caption{Full Automation of T3 in \SFiveU; Consistency of Scott's
  Axioms;  Automatic Verification of Modal Collapse} \label{Scott_S5U}
\end{figure}

@@

\subsection{Improved Embedding}\label{sec:improvedembedding}
\begin{figure}[t]
\centerline{\includegraphics[width=1\columnwidth,height=10cm]{./QML_S5U.png}}
\caption{Improved Embedding of \SFiveU} \label{QML_S5U}
\end{figure}

@@

\begin{figure}
\centerline{\includegraphics[width=1\columnwidth]{./UglyEssence.png}}
\caption{Definition of Essence using Old Syntax} \label{UglyEssence}
\end{figure}

@@

\subsection{Argument Reconstruction in Isabelle}  \label{sec:arg2}
\begin{figure}[t]
\centerline{\includegraphics[width=1\columnwidth]{./InconsistencyIsabelleK.png}}
\caption{Inconsistency of G\"odel's Axioms in HOML \K verified in Isabelle/HOL} \label{InconsistencyIsabelleK}
\end{figure}
\begin{figure*}[t]
  \centering
  \begin{subfigure}[t]{0.715\textwidth}
    \includegraphics[width=\textwidth]{./Manuscript2.png}
    \caption{G\"{o}del's manuscript, with mutually inconsistent axioms and
      definitions highlighted (with permission from the Kurt G\"odel Papers, Shelby White and Leon Levy Archives Center, Princeton, NJ, USA, on deposit at Princeton University)} \label{GoedelScript} 
  \end{subfigure}
   \begin{subfigure}[t]{0.28\textwidth}
     \includegraphics[width=\textwidth,height=7.2cm]{./Inconsistency_S5U_direct.png}
     \caption{Inconsistency in HOML \SFiveU} \label{Inconsistency_S5U} 
   \end{subfigure}
% \includegraphics[width=.73\textwidth]{./Images/Manuscript2.png} \hfill
% \includegraphics[width=.26\textwidth]{./Images/Inconsistency_S5U_direct.png}
 \caption{The inconsistency in G\"{o}del's manuscript has been
   detected and verified by HOL ATPs} 
\end{figure*}
To verify the correctness of the informal argument explained above, it
was reconstructed in Isabelle/HOL, using Metis\footnote{Metis, unlike
  external provers such as \textsc{Leo-II} or Satallax, 
  constructs proofs in Isabelle's highly trusted kernel calculus.} to automate the
inessential parts (cf. Fig.~\ref {InconsistencyIsabelleK}). The essential use of the Empty Essence Lemma, on
the other hand, is explicitly stated, to ensure that Isabelle is
reconstructing the same argument. In fact, without the help of this
lemma, Metis is still not strong enough to refute G\"odel's
axioms.
